---
bibliography: includes/bib/lit-review.bib
classoption: report
csl: includes/bib/vancouver.csl
documentclass: uncdissertation
fontsize: 12pt
link-citations: yes
output:
  pdf_document:
    includes:
      before_body: includes/tex/doc_prefix_unc.tex
      in_header: includes/tex/header-unc.tex
    keep_tex: no
    number_sections: yes
    pandoc_args:
    - --filter
    - C:/Python27/Scripts/pandoc-eqnos.exe
    toc: no
  html_document:
    toc: yes
    toc_float: yes
---


```{r include=FALSE}
library(knitr)
library(readxl)
library(htmlTable)
library(xtable)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

<!-- NOTES: 

1. All template code for formatting based on work from https://github.com/jvcasill/ua_thesis_rmd. The original .cls file is from https://github.com/mmalahe/unc-dissertation

2. Be careful of unicode characters, especially in the bibliography file. Very hard to find and can kick up errors that prevents the file from compiling. For example, a Wright reference had a U+2AB unicode character after children (an apostrophe) that was not recognized and kept the program from compiling. or 'Package inputenc Error: Unicode char α (U+3B1)', which was actually in a journal title in the .bib file. Dashes are also problematic. See header-unc.tex to avoid that problem and add characters with 'declareunicodecharacters'.

One way around this is to save the google doc file as a .txt file (ansi coding) then convert to utf-8 after opening in R (File|reopen with encoding)

Using pandoc-eqnos python program in pandoc to number equations:  https://github.com/tomduck/pandoc-eqnos
-->

<!-- Include introductory chapters -->

```{r, child='./sections/dissertation-section1.Rmd', eval=T}
```


<!-- Include first manuscript. -->
 
\chapter{Manuscript 1: Sociodemographic predictors of early postnatal growth\label{ms1}}

```{r, child='includes/scripts/paper1/draft/ms1.Rmd', eval=T}
```

<!-- Include tables and figures for first manuscript -->

```{r, child='includes/scripts/paper1/sitar-rev5/tables-ms.Rmd', eval=T}
```


<!-- Include second manuscript. -->
 
\chapter{Manuscript 2: Infant Growth Trajectories and Lipid Levels in Adolescence\: Evidence from
  a Chilean Infancy Cohort\label{ms2}}

```{r, child='includes/scripts/paper2/draft/ms2.Rmd', eval=T}
```

<!-- Include tables and figures for second manuscript -->

```{r, child='includes/scripts/paper2/bch-read-all-bf-pdfversion.Rmd', eval=T}
```

<!-- Include third manuscript. -->
 
\chapter{Manuscript 3: Infant growth as an effect modifier of genetic-lipid associations\: evidence from a Chilean infancy cohort\label{ms3}}

```{r, child='includes/scripts/paper3/draft/ms3.Rmd', eval=T}
```

<!-- Include tables and figures for third manuscript -->


```{r, child='includes/scripts/paper3/ms3-read-all-pdfversion.Rmd', eval=T}
```


<!-- Discussion section -->


```{r, child='./sections/discussion.Rmd', eval=T}
```


<!-- Include references -->

\setlength{\parindent}{-0.2in}
\setlength{\leftskip}{0.2in}
\setlength{\parskip}{8pt}

\backmatter

<!-- Include appendices -->

\appendix

```{r, child='./sections/appendix-graffar-desc.Rmd', eval=T}
```

```{r, child='./sections/appendix-model-fit-sitar.Rmd', eval=T}
```


```{r, child='./sections/appendix-model-fit-lgmm.Rmd', eval=T}
```


```{r, child='./sections/appendix-a-rev.Rmd', eval=T}
```


```{r, child='./sections/appendix-b-new.Rmd', eval=T}
```

<!--
```{r appc, child='./sections/appendix-c.Rmd', eval=T, fig.width=6, fig.height=6}
```
-->

\chapter{Background for power calculations\label{appD}}

```{r appd, child='./sections/appendix-d.Rmd', eval=T, fig.width=6, fig.height=6}
```

\chapter*{References}
\printbibliography{}
\addtocontents{toc}{\protect\vspace*{\baselineskip}}

